Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
pf <- read.delim('pseudo_facebook.tsv')
suppressMessages(library(dplyr))
library(ggplot2)
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) +
  geom_histogram(stat="identity")
```

***

### Plotting Conditional Summaries
# Write code to create a new data frame,
# called 'pf.fc_by_age_gender', that contains
# information on each age AND gender group.

# The data frame should contain the following variables:

#    mean_friend_count,
#    median_friend_count,
#    n (the number of users in each age and gender grouping)

# Here is an example of the structure of your data frame. Your
# data values will be different. Note that if you are grouping by
# more than one variable, you will probably need to call the
# ungroup() function. 

#   age gender mean_friend_count median_friend_count    n
# 1  13 female          247.2953                 150  207
# 2  13   male          184.2342                  61  265
# 3  14 female          329.1938                 245  834
# 4  14   male          157.1204                  88 1201

Notes:

```{r Plotting Conditional Summaries}
gender_group <- group_by(subset(pf, !is.na(pf$gender)), gender, age) 
# could also use the `filter` command to get rid of nas
pf.fc_by_age_gender <- summarize(gender_group,
                             mean_friend_count=mean(friend_count),
                             median_friend_count=median(friend_count),
                             n=n())
pf.fc_by_age_gender <- ungroup(pf.fc_by_age_gender)  # Summarize removed one layer of grouping, gotta remove theother one
pf.fc_by_age_gender <- arrange(pf.fc_by_age_gender, age)
head(pf.fc_by_age_gender)


## Plot of median friend count of vs age, split by gender
ggplot(aes(x=age, y=median_friend_count, color=gender), data=pf.fc_by_age_gender) + 
  geom_line()

## Similar answer if we didn't make the subset pipeline above

#ggplot(aes(x = age, y = friend_count), 
#              data = subset(pf, !is.na(gender))) +   geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:
Now we have data repeated for each year:
  #   age gender mean_friend_count median_friend_count    n
  # 1  13 female          247.2953                 150  207
  # 2  13   male          184.2342                  61  265
  # 3  14 female          329.1938                 245  834
  # 4  14   male          157.1204                  88 1201

The "Long" format will split it out so that we will have one row per year, with per-gender stats in each column
EG:
  age   male_mean_friend_count  female_mean_friend_count ...

```{r}
#install.packages('reshape2')
#install.packages("tidyr")
library(reshape2)
library(tidyr)

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, 
                                  age ~ gender,          # left of ~ is what we want to keep, right is what we want to expand
                                  value.var='median_friend_count'  # holds key values of values in new dataframe
                                  )
head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:
# Plot the ratio of the female to male median
# friend counts using the data frame
# pf.fc_by_age_gender.wide.

# Think about what geom you should use.
# Add a horizontal line to the plot with
# a y intercept of 1, which will be the
# base line. Look up the documentation
# for geom_hline to do that. Use the parameter
# linetype in geom_hline to make the
# line dashed.

# The linetype parameter can take the values 0-6:
# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash

```{r Ratio Plot}
ggplot(aes(x=age, y=female/male), data=pf.fc_by_age_gender.wide) + 
  geom_line() + 
  geom_hline(yintercept=1, alpha=0.3, linetype=2) +
  geom_smooth()
```

***

### Third Quantitative Variable
Tenure is a good additional measure for friend counts. People on FB longer have more friends

Measure with epoch being 2014 (as in zero days on FB being 2014). More tenure equals earlier years
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- 2014 - ceiling(pf$tenure/365)
```

***

### Cut a Variable
Notes:

Cut data into 4 bins: 2004-2009, 2009-2011, 2011-2012, 2012-2014
```{r Cut a Variable}
print(table(pf$year_joined))
print("Not much data for earlier years is there... cut the data!")
pf$year_joined.bucket <- cut(pf$year_joined, breaks=c(2004,2009,2011,2012,2014))
print(table(pf$year_joined.bucket))
```

***

### Plotting it All Together
Notes:
# Create a line graph of friend_count vs. age
# so that each year_joined.bucket is a line
# tracking the median user friend_count across
# age. This means you should have four different
# lines on your plot.

# You should subset the data to exclude the users
# whose year_joined.bucket is NA.

```{r Plotting it All Together}
ggplot(aes(x=age, y=friend_count), 
              data=subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color=year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x=age, y=friend_count), 
              data=subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color=year_joined.bucket), stat = 'summary', fun.y = mean) +
  geom_line(stat='summary', fun.y=mean, linetype=2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
oldies <- subset(pf, tenure >= 1)
friendrate <- oldies$friend_count/oldies$tenure
summary(friendrate)

## Better with `with` command
with(subset(pf, tenure >= 1), summary(friend_count/tenure))
```

***

### Friendships Initiated
Notes:
# Create a line graph of mean of friendships_initiated per day (of tenure)
# vs. tenure colored by year_joined.bucket.

# You need to make use of the variables tenure,
# friendships_initiated, and year_joined.bucket.

# You also need to subset the data to only consider user with at least
# one day of tenure.
 

```{r Friendships Initiated}
ggplot(aes(x=tenure, y=friendships_initiated/tenure), 
              data=subset(pf, tenure > 0)) + 
  geom_line(aes(color=year_joined.bucket), alpha=0.5, stat = 'summary', fun.y = mean) +
  geom_smooth(aes(color=year_joined.bucket))
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***
### Histograms Revisited
Notes:
Many rows per household, one per each purchase. So different from FB data

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
str(yo)

## Change the `id` from an int to a factor
yo$id <- factor(yo$id)
print(str(yo))

## histogram of yogurt prices
# qplot(yo$price)  # easy way
ggplot(aes(x=price), data=yo) + geom_histogram(binwidth = 5)
```

***

### Number of Purchases
Notes:

# Create a new variable called all.purchases,
# which gives the total counts of yogurt for
# each observation or household.

# One way to do this is using the transform
# function. You can look up the function transform
# and run the examples of code at the bottom of the
# documentation to figure out what it does.

# The transform function produces a data frame
# so if you use it then save the result to 'yo'!

# OR you can figure out another way to create the
# variable.

Households are identified by the ID, and we need to sum up all of the purchases per house over all observations
```{r Number of Purchases}
print(summary(yo$price))  # 75 percentile == maximumum, so we see the positive skew
table(yo$price)    # also see the segmented distribution of each price here in the table of prices 

yo <- transform(yo, all.purchases=strawberry+blueberry+pina.colada+plain+mixed.berry)

# verbose: yo$all.purchases = yo$strawberry + yo$blueberry + ...
```

***

### Prices over Time
Notes:

Create a scatterplot of price vs time.

We have a price column, and a time column

> From the plot, it looks like the 'modal' price increases periodicallyin time. 
The periodic lower price clusters coul represent coupons being used, sales, etc.

```{r Prices over Time}
ggplot(aes(x=time, y=price), data=yo) + 
  geom_jitter(alpha=0.3, shape=21, fill=I('#F79240'))
```

***

### Sampling Observations
Notes:

***
x %in% y returns a logical (boolean) vector the same length as x that says whether each entry in x appears in y. That is, for each entry in x, it checks to see whether it is in y. 
### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(111) # Seed random number generator
sample.ids = sample(levels(yo$id), 16)

## Plot that subsample
ggplot(aes(x=time, y=price), data=subset(yo, id %in% sample.ids)) + 
  facet_wrap(~id) + 
  geom_line() + 
  geom_point(aes(size=all.purchases), pch=1)
  
```

***
Can't always rely on domain specific knowledge. Sometimes, we want data to speak
for itself. Enter the scatterplot matrix! 

### Scatterplot Matrix
Notes:

``` {r scatter matrix}
library(GGally)

# See all fb data against eachother
set.seed(69101)
# Take subset of the columns
pf_subset = pf[, c(2:15)] # Don't want user ID, user name, or the 'bucket' varuable. 
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ])
```
### Even More Variables
Notes:

***
Genomic data has lots of info, lets heatmap it! 

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

