Lesson 6
========================================================

# Let's consider the price of a diamond and it's carat weight.
# Create a scatterplot of price (y) vs carat weight (x).

# Limit the x-axis and y-axis to omit the top 1% of values.

### Scatterplot Review

```{r Scatterplot Review}
library(ggplot2)
data(diamonds)

ggplot(aes(x=carat, y=price), data=diamonds) + 
  xlim(0,quantile(diamonds$carat, 0.99)) + 
  ylim(0,quantile(diamonds$price, 0.99)) +
  geom_point(alpha=0.3, shape=21, fill=I('#F79240')) +
  stat_smooth(method=lm)
# also seems to work??
  #scale_x_continuous(limits=c(0,quantile(diamonds$carat, 0.99)))
  #scale_y_continuous(limits=c(0,quantile(diamonds$price, 0.99)))
```

***

### Price and Carat Relationship
Response:

***

### Frances Gerety
Notes:

#### A diamonds is


***

### The Rise of Diamonds
Notes:

***

### ggpairs Function
Notes:

```{r ggpairs Function}
# install these if necessary
# install.packages('GGally')
# install.packages('scales')
# install.packages('memisc')
# install.packages('lattice')
# install.packages('MASS')
# install.packages('car')
# install.packages('reshape')
# install.packages('plyr')

# load the ggplot graphics package and the others
library(ggplot2)
library(GGally)
library(scales)
library(memisc)

# sample 10,000 diamonds from the data set
set.seed(20022012)
diamond_samp <- diamonds[sample(1:length(diamonds$price), 10000), ]
ggpairs(diamond_samp, 
        lower = list(continuous = wrap("points", shape = I('.'))), 
        upper = list(combo = wrap("box", outlier.shape = I('.'))))
```

## Log scale because price is skewed!
```{r}
library(gridExtra)
plot1 <- qplot(x=price, data=diamonds, binwidth=200, fill=I("#099DD9")) + 
  ggtitle('Price')
plot2 <- qplot(x=price, data=diamonds, binwidth=0.02, fill=I("#F79420")) + 
  scale_x_log10() +
  ggtitle('Price (log10)')
grid.arrange(plot1,plot2,ncol=2)
```

### The Demand of Diamonds
Notes:
Plot price on log10 scale. Price is pretty skewed so we want to do a transformation

```{r The Demand of Diamonds}
ggplot(aes(x=carat, y=price), data=diamonds) + 
  scale_y_log10() + 
  geom_point() +
  ggtitle("Log10 price by carat")
```

***

### Connecting Demand and Price Distributions
Notes:

We think that we can do better. Since:
 * price is a function of carat
 * Carat is a function of volume
 * Volume is cubic (X x Y x Z), we want to cube root carat

### Create a new function to transform the carat variable

```{r cuberoot transformation}
cuberoot_trans = function() trans_new('cuberoot', transform = function(x) x^(1/3),
                                      inverse = function(x) x^3)
```

#### Use the cuberoot_trans function
```{r Use cuberoot_trans}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point() + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  geom_smooth(method='lm') + 
  ggtitle('Price (log10) by Cube-Root of Carat')
print("Hey it looks pretty darn linear now! How useful is this though...?")
```

***

### Overplotting Revisited
Let's look at the top prices and carat values in a table

```{r Sort and Head Tables}
print(head(sort(table(diamonds$carat), decreasing=TRUE)))
print(head(sort(table(diamonds$price), decreasing=TRUE)))
print("WOAH! Look at all those identical values (2nd row shows many counts for each price/carat value)")
```

## Overplotting
We have significent amount of identical points. Hence mega overplotting. Let's work out
a way to have less overplotting and visualize this better

```{r Overplotting Revisited}
# diamonds <- aggregate(diamonds$price,by=list(x=diamonds$price,y=diamonds$carat),length)
# names(diamonds[-1]) <- "count"

ggplot(aes(carat, price), data = diamonds) + 
  geom_point(alpha=0.05, 
             position=position_jitter(height=0, width=0), shape=21,
             fill=I("#F79420")) + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
 # Alternately, by using size as a density indicator instead of alpha...
  # geom_count(position=position_jitter(height=0, width=0), shape=21,
  #            fill=I("#F79420"))
```

***

### Other Qualitative Factors
Notes:
# Adjust the code below to color the points by clarity.

# A layer called scale_color_brewer() has 
# been added to adjust the legend and
# provide custom colors.

# See if you can figure out what it does.
# Links to resources are in the Instructor Notes.

# You will need to install the package RColorBrewer
# in R to get the same colors and color palettes.

Alter the code below.
```{r Price vs. Carat and Clarity}
# install and load the RColorBrewer package
# install.packages('RColorBrewer')
library(RColorBrewer)

ggplot(aes(x = carat, y = price, color=clarity), data = diamonds) + # <--- this is the part that recolors just in the aes...
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```

***

### Clarity and Price
Response:

***

### Price vs. Carat and Cut

Alter the code below.
```{r Price vs. Carat and Cut}
ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Cut', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Cut')
```

***

### Cut and Price
Response:

***

### Price vs. Carat and Color

Alter the code below.
```{r Price vs. Carat and Color}
ggplot(aes(x = carat, y = price, color = color), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = "Color\nCode",
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Coilor')
```

***

### Color and Price
Response:

***

### Linear Models in R
Notes:

Response:

***

### Building the Linear Model
NEAT! 

Reference : http://data.princeton.edu/R/linearModels.html

```{r Building the Linear Model}
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamonds)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5)
```

Notice how adding cut to our model does not help explain much of the variance
in the price of diamonds. This fits with out exploration earlier.

***

### Model Problems
Video Notes:

Research:
(Take some time to come up with 2-4 problems for the model)
(You should 10-20 min on this)

Response:

***

### A Bigger, Better Data Set
Notes:

```{r A Bigger, Better Data Set}
# install.packages('bitops')
# install.packages('RCurl')
library('bitops')
library('RCurl')

diamondsurl = getBinaryURL("https://raw.github.com/solomonm/diamonds-data/master/BigDiamonds.Rda")
# load(rawConnection(diamondsurl))
load("BigDiamonds.Rda")
```

The code used to obtain the data is available here:
https://github.com/solomonm/diamonds-data

## Building a Model Using the Big Diamonds Data Set
Notes:

**Use only a subset**

```{r Building a Model Using the Big Diamonds Data Set}
diamondsbig <- subset(diamondsbig, !is.na(diamondsbig$price))
diamondsbig <- subset(diamondsbig, diamondsbig$price<quantile(diamondsbig$price, 0.99, na.rm=TRUE))
## Take random sample
set.seed(20022012)
bigdiamond_samp <- diamondsbig[sample(1:length(diamondsbig$price), 10000), ]

m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = bigdiamond_samp)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5)

```


***

## Predictions

Example Diamond from BlueNile:
Round 1.00 Very Good I VS1 $5,601

```{r}
#Be sure you’ve loaded the library memisc and have m5 saved as an object in your workspace.
thisDiamond = data.frame(carat = 1.00, cut = "V.Good",
                         color = "I", clarity="VS1")
modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)
print(modelEstimate)       # Model prediction is the LOG of price: (log(price) ~ carat^1/3)
print(exp(modelEstimate))
```

Evaluate how well the model predicts the BlueNile diamond's price. Think about the fitted point estimate as well as the 95% CI.

***

## Final Thoughts
Notes:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

