Lesson 4
========================================================
```{r startingout}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep='\t')
head(pf)
```
***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
qplot(x=age, y=friend_count, data=pf)
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
# GGPlot syntax is more verbose and more powerful. Declare layers and whatnot
# qplot version: qplot(x=age, y=friend_count, data=pf)

# Specify while adding one layer at a time
ggplot(aes(x=age, y=friend_count), data=pf) +
  #geom_point(alpha=1/20) + # This means it takes 20 points on top of eachother to be fully darkened. Neat
  geom_jitter(alpha=1/20) + # Similar to geom_point (Scatter) but adds jitter since `age` is actually continuous but takes on only integer values for us.
  xlim(13,90)

## Summary stats
summary(pf$age)
```

***

### Coord_trans()
Notes:

```{r Coord_trans()}

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_point(alpha=1/20, position=position_jitter(h=0)) + # Saying we want to jitter the height with minimum height of 0. Avoids errors with sqrt of negative numbers
  xlim(13,90) + 
  coord_trans(x="identity", y="sqrt")
```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x=age, y=friendships_initiated), data=pf) + 
  geom_point(alpha=1/5, position=position_jitter(h=0)) + 
  xlim(13,90) + 
  coord_trans(x="identity", y="sqrt")
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
# Library to split up dataframe and apply function to parts of data 
# install.packages('dplyr')
library(dplyr)

# filter() groupby(), etc
age_groups <- group_by(pf, age)
pf.fc_by_age <- summarize(age_groups,
          friend_count_mean=mean(friend_count),
          friend_count_median=median(friend_count),
          n=n()) # Get number of users in each group. `n()` only works in summarize

# Make sure ages go in order
pf.fc_by_age <- arrange(pf.fc_by_age, age)
head(pf.fc_by_age)

```
```{r Alternative Conditional Means}
# Library to split up dataframe and apply function to parts of data 
library(dplyr)

pf.fc_by_age2 <- pf %>%                # %.% lets you chain functions onto your dataframe
  group_by(age) %>%
  summarize(friend_count_mean=mean(friend_count),
          friend_count_median=median(friend_count),
          n=n()) %>%
  arrange(age)

head(pf.fc_by_age2)
```
Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x=age, y=friend_count_mean), data=pf.fc_by_age) +
  geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
# Original plot with color changed
ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_point(alpha=1/20, position=position_jitter(h=0), color='orange') + # Saying we want to jitter the height with minimum height of 0. Avoids errors with sqrt of negative numbers
  xlim(13,90) + 
  coord_trans(x="identity", y="sqrt") +
  geom_line(stat='summary', fun.y=mean) + # Add a summary line showing mean friend count
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.1), linetype=2, color='blue') + # Add a sumamry showing 10% quantile
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.9), linetype=2, color='blue') + # 90th  percentile
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.5), color='blue')               # Median
  #coord_cartesian(xlim=c(50,90))
```

### Correlation
Notes:

Pearson's correlation test `r` 
* https://s3.amazonaws.com/udacity-hosted-downloads/ud651/correlation_images.jpeg
* http://www.r-tutor.com/elementary-statistics/numerical-measures/correlation-coefficient
* **Correlation function in R** http://www.statmethods.net/stats/correlations.html

```{r Correlation}
# Same as `cor()`, but with better confidence intervals values provided
cor.test(pf$age, pf$friend_count)

## Alternately, use the `with` function around a dataframe to evaluate an R
# expression in an environment 'constructed from the data'
with(pf, cor.test(age, friend_count, method='pearson'))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
-0.027
Response:

***

### Correlation on Subsets
Notes:
Old people screw up the stats here. So just subeset them out of consideration

```{r Correlation on Subsets}
with(subset(pf, age<= 70 & age >= 13), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

Can try `cor.test(..., method='spearman')` for other possible correlation measures
http://www.statisticssolutions.com/correlation-pearson-kendall-spearman/ Kindof a shitty explanation

***

## Create Scatterplots
Should be highly correlated
Notes:

```{r likes vs online likes}
ggplot(aes(x=www_likes_received, y=likes_received), data=pf) + 
  geom_point(alpha=1/20, position=position_jitter(w=0,h=0), color='green') + # Saying we want to jitter the height with minimum height of 0. Avoids errors with sqrt of negative numbers
  xlim(0,quantile(pf$www_likes_received, 0.95)) +  # Take 95th percentile values only to get rid of super outliers
  ylim(0,quantile(pf$www_likes_received, 0.95)) +
  #coord_trans(x="sqrt", y="sqrt") +
  geom_line(stat='summary', fun.y=mean) + # Add a summary line showing mean friend count
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.1), linetype=2, color='blue') + # Add a sumamry showing 10% quantile
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.9), linetype=2, color='blue') + # 90th  percentile
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.5), color='blue')  +           # Median
  geom_smooth(method='lm', color='red')                                           # Dope! Add a `linear model`

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
cor.test(pf$www_likes_received, pf$likes_received)
```

=

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
data(Mitchell)
# Fugly
ggplot(aes(x=Month, y=Temp), data=Mitchell) + 
  geom_point()
print(cor.test(Mitchell$Month, Mitchell$Temp))
```

Create your plot!

```{r Temp vs Month}
# Group months together. Just modulo month number with 12!
ggplot(aes(x=Month%%12, y=Temp), data=Mitchell) + 
  geom_point() +
  # Put `Month` into discrete units since it really isn't 'continuous'
  scale_x_continuous( breaks=seq(0,range(Mitchell)[2], 12))
```


***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(aes(x=age, y=friend_count_mean), data=pf.fc_by_age) + 
  geom_line()
  
pf$age_with_months <- pf$age + ((12-pf$dob_month) / 12)

```

***

### Age with Months Means
# Create a new data frame called
# pf.fc_by_age_months that contains
# the mean friend count, the median friend
# count, and the number of users in each
# group of age_with_months. The rows of the
# data framed should be arranged in increasing
# order by the age_with_months variable.

# For example, the first two rows of the resulting
# data frame would look something like...

# age_with_months  friend_count_mean	friend_count_median	n
#              13            275.0000                   275 2
#        13.25000            133.2000                   101 11



```{r Age with Months Means}
library(dplyr)

pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarize(friend_count_mean=mean(friend_count),
          friend_count_median=median(friend_count),
          n=n()) %>%
  arrange(age_with_months)

head(pf.fc_by_age_months)
```

Programming Assignment - plot with ages < 71
```{r Programming Assignment}
ggplot(aes(x=age_with_months, y=friend_count_mean), 
       data=subset(pf.fc_by_age_months, age_with_months<71)) +
  geom_line() +
  #coord_trans(x="sqrt", y="sqrt") +
  geom_smooth(method='auto', color='red')  
```


Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

