Lesson 3
========================================================

***

### What to Do First?
Notes: Download data from: https://s3.amazonaws.com/udacity-hosted-downloads/ud651/pseudo_facebook.tsv

***

### Pseudo-Facebook User Data
Notes: Read it in

```{r Pseudo-Facebook User Data}
getwd()
pf <- read.csv('pseudo_facebook.tsv', sep='\t')
```

***

### Histogram of Users' Birthdays
Notes: GGplot is nicer than the base packages library

```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)

install.packages('ggthemes', dependencies = TRUE) 
library(ggthemes) 

theme_set(theme_minimal(24)) 
```

```{r}
# Plot command - WTF why doesn't this work?? Simple as can be and follows instructor notes
qplot(x=dob_day, data=pf) +
  scale_x_continuous(breaks=1:31) + 
  facet_wrap(~dob_month, ncol=3)   # Formula for what we split facets over
```

***

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}

```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
# qplot(x=friend_count, data=pf, xlim=c(0,1000), binwidth=25) +  # c(0,1000) is a continuous array from 0-1000
#   scale_x_continuous(limits=c(0,1000), breaks=seq(0, 1000, 50))


## Equivalent with a 'layer'
qplot(x=friend_count, data=subset(pf, !is.na(gender)), binwidth=10) +
  scale_x_continuous(limits=c(0,1000), breaks=seq(0, 1000, 100)) + 
  facet_grid(~gender)  # Need to omit N/A gender values
```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}

```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```



### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)  # tabulate gender counts

## `by` takes <data, indices, function to apply>
#   | variable       | categorical subset   | `summary` is the function
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

```{r Tenure}
qplot(x=tenure, data=pf, binwidth=30, color=I('black'), fill=I('#099DD9'))
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
#     | on the fly convert DAYS of tenure to YEARS of tenure
qplot(x=tenure/365, data=pf, binwidth=.25, color=I('black'), fill=I('#F79420'),
      xlab="Tenure in years on Facebook") +   # Can add with kwarg 
  scale_x_continuous(breaks=seq(1,7,1), limits=c(0,7)) + 
  ylab("Count of Users")                      # Or with Layer
```

### User Ages
Notes:

```{r User Ages}
qplot(x=age, data=pf, fill=I('gold'), color=I('Black'), binwidth=1) + 
  scale_x_continuous(breaks=seq(10,120,10), limits=c(10,120))
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
```{r transform}
print(summary(pf$friend_count))
print(summary(log10(pf$friend_count+1)))  # add 1 to avoid 0 limit issues
print(summary(sqrt(pf$friend_count+1)))

#install.packages("gridExtra")
library(gridExtra)

p1=qplot(x=friend_count, data=pf)
p2=qplot(x=log10(friend_count+1), data=pf)
p3=qplot(x=sqrt(friend_count+1), data=pf)
grid.arrange(p1,p2,p3)
```

### Transform with ggplot syntax alternate syntax

```{r Add a Scaling Layer}
p1 <- ggplot(aes(x=friend_count), data=pf) + geom_histogram()  # Base plot directive
p2 <- p1 + scale_x_log10()  # Note that with this layer method, axis labels actually reflect true values
p3 <- p1 + scale_x_sqrt()
grid.arrange(p1,p2,p3)
```

***


### Frequency Polygons

```{r Frequency Polygons}
## Original histogram of friend count vs gender
qplot(x=friend_count, data=subset(pf, !is.na(gender)), binwidth=10) +
  scale_x_continuous(lim=c(0,1000), breaks=seq(0,1000,100)) + 
  facet_wrap(~gender)

## New overlaid polygon method
qplot(x=friend_count, data=subset(pf, !is.na(gender)),
      binwidth=10, geom='freqpoly', color=gender) +
  scale_x_continuous(lim=c(0,1000), breaks=seq(0,1000,100))

## Change it up so 'y' now shows proportions instead of just raw counts
qplot(x=friend_count, y=..count../sum(..count..),
      data=subset(pf, !is.na(gender)),
      binwidth=10, geom='freqpoly', color=gender) +
  scale_x_continuous(lim=c(0,1000), breaks=seq(0,1000,100)) +
  xlab('Friend Count') + 
  ylab('Proportion of Users with that Friend Count')
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
qplot(x=www_likes,data=subset(pf, !is.na(gender)),
      binwidth=1, geom='freqpoly', color=gender) +
  scale_x_continuous() + 
  scale_x_log10() +
  xlab('Likes on the web') + 
  ylab('Proportion of Users with that Friend Count')

# stats
by(pf$www_likes, pf$gender, sum)
```


***

### Box Plots
Notes:

```{r Box Plots}
## X axis is categorical (grouped), Y axis is descriptive statistic (continuous)
qplot(x=gender, y=friend_count, 
      data=subset(pf, !is.na(gender)), 
      geom='boxplot')
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
qplot(x=gender, y=friend_count, 
      data=subset(pf, !is.na(gender)), 
      geom='boxplot') + 
  #ylim(c(0,1000)) +   #or 
  #scale_y_continuous(limits=c(0,1000)) # But these two methods THROW AWAY data before calculating and drawing...
  coord_cartesian(ylim=c(0,1000))  # This method is nice because it just limits the cartesian coordinates drawn
```
```{r}
## Let's look at actual friend_count summarized vs gender 
by(pf$friend_count, pf$gender, summary)
```
***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
qplot(x=gender, y=friendships_initiated, 
      data=subset(pf, !is.na(gender)), 
      geom='boxplot') + 
  coord_cartesian(ylim=c(0,500))  # This method is nice because it just limits the cartesian coordinates drawn
```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}

by(pf$friendships_initiated, pf$gender, median)
```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
summary(pf$mobile_likes)
# versus...
summary(pf$mobile_likes > 0)

## Create a new variable
mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0 )  # 1 if mobile_likes > 0, 0 otherwise
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)
length(subset(pf$mobile_check_in, pf$mobile_check_in==1))/length(pf$mobile_check_in)
```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!