Vines and Wines by Zach Dischner
========================================================

**Date:** April 4 2017
Part of the Udacity Data Analyst Nanodegree program

![Red Wines](https://goo.gl/images/uynJff)



```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load up all packages we want to use
library(ggplot2)
library(gridExtra)
library(GGally)
library(scales)
library(memisc)
```

```{r echo=FALSE, Load_the_Data}
setwd("/Users/dischnerz/code/udatascience/projects/p4-R-investigation")
# Load the Data
data <- read.csv("wineQualityReds.csv")

## Some preliminary manipulation
data$total.acidity <- data$fixed.acidity + data$volatile.acidity
```

### Dataset Description
This investigaion is about red wine.

```{r echo=FALSE}
str(data)
```
# Univariate Plots Section
Overall distribution of *quality* rating of red wines. Quality here is a single
measurement between 1 and 10 of oerall wine quality, as decided by three 
separate agencies

```{r echo=FALSE, Univariate_Plots}
# Easiest, look at quality
qplot(x=quality, data=data, geom='bar')
```

Overall wine quality, rated on a scale from 1 to 10, has a normal shape and
very few exceptionally high or low quality ratings. 

```{r echo=FALSE , stats}
# Fixed and Volatile acidity
q1<-ggplot(aes(x=pH), data=data)+
   geom_histogram(color =I('black'),fill = I('#990000'))+
  ggtitle('pH distribution')
q2<-ggplot(aes(x=sulphates), data=data)+
   geom_histogram(color =I('black'),fill = I('#990000'))+
  ggtitle('Sulphates distribution')
q3<-ggplot(aes(x=chlorides), data=data)+
   geom_histogram(color =I('black'),fill = I('#990000'))+
  ggtitle('Chlorides distribution')
q4<-ggplot(aes(x=citric.acid), data=data)+
   geom_histogram(color =I('black'),fill = I('#990000'))+
  ggtitle('Citric Acid')
q5<-ggplot(aes(x=total.sulfur.dioxide), data=data)+
   geom_histogram(color =I('black'),fill = I('#990000'))+
  ggtitle('Total SO2 distribution')
q6<-ggplot(aes(x=total.acidity), data=data)+
   geom_histogram(color =I('black'),fill = I('#990000'))+
  ggtitle('Total Acidity (fixed + volatile)')

grid.arrange(q1,q2,q3,q4,q5,q6,ncol=2)
```
Of all single component distributions, `pH` appears to be the only measurement 
that is normally distributed. `Acidity` is mostly normal with just a slight 
positive skew. `Citric Acid` and `S02` particularly exhibit a strong positive 
skew. `pH` and `Acidity` reflect measurements that are centered around non-zero
numbers, whereas the other measurements are clustered at near-zero numbers. 
Since these measurements cannot have negative values, the skewed distribution is
sensible. 

**Note on Acidity**
Wine acidity is measured in multiple components: volatile and static. Total 
acidity is not to be confused with the `pH` of a wine. The former details the 
amount of acid in the wine, the latter relates to the *strength* of the acids. 

More on *static* and *volatile* acidity measures here: 
http://winemakersacademy.com/understanding-wine-acidity/

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

#### Overview
The red wine dataset features 1599 separate observations, each for a different 
red wine sample. For each sample, 11 *laboratory* based measurements were made,
as well as a single subjective overall quality rating. The quality of a wine was
decided upon by three separate professional wine assessment institutions. 

As presented, each wine sample is provided as a single row in the dataset. Due to
the nature of how some measurements are gathered, some values given represent 
*components* of a measurement total. For example, `data.fixed.acidity` and 
`data.volatile.acidity` are both obtained via separate measurement techniques, 
and must be summed to indicate the total acidity present in a wine sample. For 
these cases, I supplimented the data given by computing the total and storing in 
the dataframe with a `data.total.*` variable. 

#### Features of Interest
The main interesting measurement here is the wine `quality`. It is the
subjective measurement of how attractive the wine might be to a consumer. The
goal here will be to try and correlate non-subjective wine properties with its 
quality.

I am curious about a few trends in particular:
* sulphates vs quality - Low sulphate wine has a reputation for not causing \
hangovers
* alcohol vs quality - Just an interesting measurement. Strong beers are \
typically higher quality and harder to make than lower proof beers. 

At first, the lack of an age metric might be surprising since it is commonly
a factor in quick assumptions of wine quality. However, since the actual effect
of wine age is on the wine's measurable chemical properties, its inclusion here
is not necessary. 

### Distributions
I left distributions alone here. Many measurements that were clustered close to
zero had a positive skew (you can't have negative percentages or ammounts).
Others such as `pH` and `total.acidity` and `quality` had normal looking 
distributions.


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}
# sample 10,000 diamonds from the data set
set.seed(20022012)
diamond_samp <- diamonds[sample(1:length(diamonds$price), 10000), ]
ggpairs(diamond_samp, 
        lower = list(continuous = wrap("points", shape = I('.'))), 
        upper = list(combo = wrap("box", outlier.shape = I('.'))))
```

# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!